---
title: "Epi1"
author: "Rasmus Larsen"
date: "`r Sys.Date()`"
output:
  html_document:
    css: null
    theme: null
vignette: >
  %\VignetteIndexEntry{epi_vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(epi1)
library(dplyr)
```


The Epi1 package creates a frequency and percentage table commonly used in
epidemiological papers. It is a wrapper for the [htmlTables-package](https://cran.r-project.org/web/packages/htmlTable/vignettes/general.html).

```{r}
head(example_data)
```

Counts and frequencies of the variables `color`and `clarity` grouped by `cut` is obtained below:
```{r, echo=TRUE, results='asis'}
freq_by(dataset = example_data, var_vector = c("color", "clarity"), by_group = "cut" )
```

The `by_group` variable must be of class factor, to ensure correct order.  



To do in a possible future:

* Add totals to each of the `by_group`s
* option to specify better names to the rgroup, not just the variable names.
* function for adding reference levels to regression models.
* include other summary statistics
* option to render blank cells containing few observations, to ensure anonymity.

The output is a HTML-table created by the htmlTable-package. The HTML-output 
looks best if no CSS is used.







